Transcriptomic Analysis for Prognostic Value in Head and Neck Squamous Cell Carcinoma

### 2021-07-01 main.tex
Updates from Overleaf

%%% ***
In summary, identifying predictive biomarkers for selecting standard-of-care or advanced systemic therapy\cite{Cristina2019} in \acrshort{hnscc} is crucial.

\begin{MyColorPar}{red} %%% ***
There are three essential points of biomarker discovery from survival analysis of the TCGA HNSCC dataset.

% risk of *** Answer 2-3 併入此
Second, trying to find an optimal cutpoint of that \acrshort{rna} expression data to maximize candidate mining coverage, this strategy could identify more but sometimes weak "biomarkers".
We have checked this issue in detail (in the Discussions section).
Thus, the false positives and adjustment of multiple comparisons should be corrected by \acrfull{fdr}.
Furthermore, validation of those candidates is required in the other independent dataset.
% in larger patient populations.% in larger patient
%warrant.

% dissection
%ZSWIM2\_archive.Rda % 其實最困難的在於 unknown error (ZSWIM2) 
Third, we analyzed the error log during the cutoff finding and Cox modeling, 
the result shows that program running could be halted by several technical problems.
These include:
\begin{outline}

Our approach describes a workflow implemented in the R script, which runs on the Rstudio server.
Its function includes data retrieving and pre-processing, feature selection, sliding-window cutoff mining, Kaplan-Meier survival analysis, Cox proportional hazard modeling, and biomarker discovery.
The independent HNSCC datasets were used to validate this strategy.
The workflow, shown in Figure \ref{fig:figure1}, scanned 20,500 human protein-coding genes of the \acrshort{tcga} \acrshort{hnscc} cohort to yield a model with biomarker estimate.

@@

In summary, those three biomarker candidates , clinical T stage, and surgical margin are independent prognosis factors in \acrshort{hnscc}.
We also found those candidates have proper effect size---the Cox's HR either $> 2.0$ or $< 0.5$.
Thus, the prognosis model with coefficients is established from \acrshort{tcga} and GSE65858 \acrshort{hnscc} cohorts. 
%The important input $X_1...X_n$ should be patients' features: age, specific ten gene expressions, clinical T stage, and surgical margin.
%\\[0.5cm]
% total 6+4 =10
% After validation with an independent \acrshort{hnscc} cohort (GSE2837), the result showed six overexpressed genes (symbol as CAMK2N1, PGK1, SURF4, USP10, NDFIP1, FOXA2) are significantly associated with a poor prognosis of overall survival. 
%Furthermore, the four overexpressed genes (symbol as IL19, FCGBP, IQCN - former symbol as KIAA1683, and NPB) are correlated with better survival.

@@

%%% Statistical significance (\textit{P} value) is affected by both sample size, error, effect size (substantive significance)\cite{Sullivan2012}\cite{Thiese2016}, and cutoff.
\subsection{\textit{P} value is affected by sample size and cutoff}
Statistical significance (\textit{P} value) is affected by both sample size, error, effect size (substantive significance)\cite{Sullivan2012}\cite{Thiese2016}, and cutoff. 
The effect size is the magnitude of the difference (e.g., \acrlong{hr}) among comparing groups.
%A \textit{P} value is a tool to help interpret findings from biomedical research.
%In reporting and interpreting studies, both 
%the substantive significance (effect size) and statistical significance (P value) are essential results to be reported.
While statistical significance can inform the researcher whether an effect exists, 
the \textit{P} value will not directly tell the effect size.
%The effect size is the main finding of a quantitative study.
%Statistical significance depends upon both sample size and effect size. 
The effect size is independent of the sample size\cite{Sullivan2012}.

%When we increase the sample size, the standard error decrease
%, or increase the difference between the sample statistic and hypothesized parameter, the p value decreases, thus making it more likely that we reject the null hypothesis.
In a study with large sample size, the difference could be noticed easily (i.e., \textit{P} value $< 0.05$) due to a decreased standard error\cite{Sullivan2012}.
However, the small effect size (non-zero) is often meaningless or substantive insignificance.
%With a sufficiently large sample, a statistical test will always reveal a significant difference, unless the effect size is exactly zero.
%, that is, when the effect size is exactly zero; 
%yet very small differences, even if significant, are often meaningless. 
%Thus, reporting only the significant P value for an analysis is not adequate for readers to fully understand the results.
%If the magnitude of effect is small and clinically unimportant, the p-value can be "significant" if the sample size is large.
Conversely, the effect size can be large but fail to get statistical significance if the sample size is small. 
%In this scenario, the \textit{P} value tells the effect does not exist.

The following error could also impact the \textit{P} value:
\begin{outline}
\1 A random error, defined as the variability in data, is not considered a bias but rather occurs randomly across the entire study population and can distort the measurement process (e.g., RNA-Seq experiments).
The larger sample size could reduce the random error. 
\1 A systematic error is a bias, which includes selection bias, information bias, and confounding.
It could deleteriously impact the statistical significance.
The larger sample size could not affect the systematic error.
\end{outline}
% effect size is large (whatever is true or artificial), the p value is small
Thus, if there is no error in two study groups, and the sample size is the same, the group, which has a larger effect size, will has a small \textit{P} value\cite{Thiese2016}.
If a skewed cutoff that split, for example, 425 vs. 75 between the two groups, it will also get statistical significance by increasing effect size artificially.
% https://stats.stackexchange.com/questions/162359/how-can-you-have-p-0-0001-with-a-sample-size-of-89-and-a-control-group-of-52
%As you can see, even with only 12 events (fustat = 1) that split 10 vs 2 between the two dummy groups you get a log-rank p-value < 0.0001.
%effect size, sample size, cutoff 影響 p value!

@@

We will explain the above mentioned by an example.
%***NDFIP1  % 給他們一個機會
Reviewing the special case of genes, such as NDFIP1, DKC1, PNMA5, and NPB,
we have noticed that NDFIP1, with a \textit{P} value of 0.05 at 50\% quantile (median) cutoff, could achieve a \textit{P} value of \num{2.62e-6} at 70\% quantile cutoff.
NDFIP1 has the FDR-corrected \textit{P} value as \num[round-precision=3, round-mode=figures,
scientific-notation=true]{0.0001074233} (please see supplementary figure S1, a "W"-shaped \textit{P} value plot). % it is \ref{fig:pvaluePlot_NDFIP1
%Moreover, the gene NDFIP1, one of our 20 preliminary candidates, has a \textit{P} value (around $0.05$) at 50\% quantile cutoff, achieving a \textit{P} value of \num{2.62e-6} at 70\% quantile cutoff (please see Figure \ref{fig:pvaluePlot_NDFIP1}, and also in supplementary figure S1).
%After the FDR correction, NDFIP1 still has \textit{P} value of \num[round-precision=3, round-mode=figures, scientific-notation=true]{0.0001074233}.
% the effect size
Then it has been excluded as a candidate by small effect size (HR = 1.33 in GSE65858 cohort and HR = 1.42 in GSE2837 cohort) less than 1.8.
%%%
\end{MyColorPar} % red

\subsection{Validation by Web-based Tools}
% GSE2837 validated candidates: 10
We also utilized another approach of web-based tools (TCGA) to analyze our candidates.
%\subsubsection*{SurvExpress} % TCGA
% 2021/01/23 pdf x 12 from SurvExpress
SurvExpress\cite{Aguirre-Gamboa2013} is one of the web-based tools for Biomarker comparison and validation of survival gene expression data. Using their TCGA HNSCC cohort (June 2016, n=502), the survival significance (Kaplan Meier \textit{P} value) shows in 6 genes:\\ % SurvExpress 有類似的結果:
(1) poor prognosis: CAMK2N1 (0.000006), PGK1 (0.01), SURF4 (0.002), USP10 (0.002), NDFIP1 (0.000008);\\
%, STIP1 (0.001), DKC1 (0.01);\\
%DKK1 (0.00005), CAMK2N1 (0.048214), STC2 (0.001), PGK1 (0.009978), SURF4 (0.023127), USP10 (0.017768), NDFIP1 (0.022758), FOXA2 (0.001587); % FOXA2 (0.038125)
(2) better prognosis: FCGBP (0.001).\\
%ZNF557 (0.0007), ZNF266 (0.00008), IL19 (0.049731), FCGBP (0.005658), KIAA1683 (IQCN, 0.005886), NPB (0.014177);
%17 out of 20 candidates
%?DKK1 (0.253635),
The 6 out of 10 candidate genes are also confirmed by a comparative study using SurvExpress with the TCGA cohort. Please see Supplementary Figure S2.%\ref{fig:fig_SurvExpress}.
% moved to supplementary file: Supplementary_figures.tex
% \linksupplementary{Supplementary_figures.pdf},
% % SurvExpress 沒有類似的結果是因為:
%Nevertheless, the survival prediction could not be found in FOXA2, IL19, MYO1H, LOC148709, EVPLL, PNMA5, IQCN, KIAA1683, NPB using SurvExpress. 
Nevertheless, the survival prediction could not be found in FOXA2, IL19, IQCN, KIAA1683, or NPB using SurvExpress.
Our workflow has the advantage to find an optimal cutpoint of that \acrshort{rna} expression data to maximize coverage during candidate mining.
%因為 cutoff 不適當，這也就是我們 pvalueTex \textit{P} 的設計目的與強項

%\subsubsection{HPA} %still TCGA's RNA-Seq
The Human Protein Atlas project (HPA) has proteomics analysis based on 26,941 antibodies targeting 17,165 unique proteins. The HPA's Pathology Atlas analyzes each protein in patients, using \acrshort{ihc} analysis based on tissue microarrays (TMAs) adopted from TCGA. Kaplan-Meier survival analyses are based on RNA-Seq expression levels of human genes in HNSCC tissue and the clinical outcome.
All transcriptomics data has been retrieved from the TCGA, and all proteomics data has been generated in-house using the same antibodies.
Our 5 candidates (e.g. CAMK2N1, PGK1, SURF4, USP10, and NDFIP1) are also on the list of unfavorable prognostic genes for HNSCC from \acrshort{hpa} (available at https://www.proteinatlas.org/humanproteome/pathology/head+and+neck+cancer, Version: 20.0 updated: 2020-11-19). 
FCGBP is on the list of favorable prognostic genes as well.
% DKK1, CAMK2N1, STC2, PGK1, SURF4, USP10, NDFIP1, STIP1, and DKC1
% ZNF557, ZNF266, and FCGBP
Moreover, overexpression of ubiquitin-specific peptidase 10 (USP10), one of the autophagy-related genes, has a poor prognostic impact on HNSCC, which has been proved at mRNA and protein level (HPA database, available at https://www.proteinatlas.\\
org/ENSG00000103194-USP10/pathology/head+and+neck+cancer)\cite{Ren2020}.
% 13 ARGs (GABARAPL1, ITGA3, USP10, ST13, MAPK9, PRKN, FADD, IKBKB, ITPR1, TP73, MAP2K7, CDKN2A, and EEF2K) with prognostic value were identified in HNSCC patients
%found at Human Protein Atlas dataset 
(Please see Supplementary Figure S3)%\ref{fig:fig_HPA_USP10})
% moved to Supplementary

@@

%%%
\subsection{The Candidates Shown in Literature Review}

%%
%\subsection*{Review articles}
Since the discovery of workflow, we performed the literature review by Embase/Pubmed.% to find the evidence convincing the suggested biomarkers in cancer research. 
The 5 out of 10 candidates has also been suggested by published studies using TCGA cohort, in-house cohort, or experiments in vitro and in vivo.

% PGK1
Phosphoglycerate kinase 1 (PGK1 or PGK-1), a glycolysis enzyme, is responsive in cisplatin-resistant HNSCC cell line (H-1R). 
The resistance is associated with the up-regulated expression of PGK1, CD55, ATP-binding cassette transporter genes (e.g., MDR1, MRP1, and MRP2), and down-regulated expression Caveolin 1\cite{Nakamura2005}.
PGK1 is responsive in hypoxia status, since it is one of \acrshort{hif}-1 alpha regulated gene revealed in HNSCC\cite{Peng2011}\cite{Shih2017}.

%SURF4
The patients with tumors (such as glioma, breast cancer, lymphoma, pancreatic cancer, adrenocortical carcinoma, and sarcoma%; not HNSCC
) exhibiting high Surfeit gene 4 (SURF4) expression have significantly shorter overall survival than low SURF4 expression.
SURF4 can induce cellular transformation and cell migration in vitro and has increased tumor growth ability in vivo\cite{Kim2018a}.
%NIH/3T3 cells (Mus musculus, mouse), 293T cell
%Loss of contact inhibition leads to phenotypic changes and promotes foci formation in vitro
%http://www.canevolve.org/AnalysisResults/AnalysisResults.html
%web-based tools survival (not TCGA?)

% USP10
Ubiquitin specific peptidase 10 (USP10), one of the autophagy-related genes (ARGs), may serve as a prognostic biomarker and target for various human cancers, including epithelial ovarian cancer, colorectal cancer, and HNSCC\cite{Ren2020}.
Functional annotation - by gene set variation analysis (GSVA) and gene set enrichment analysis (GSEA) - reveals that USP10 is significantly enriched in many critical pathways correlated with tumorigenesis of HNSCC, including the p53 pathway, IL-2–STAT5 signaling, transforming growth factor beta (TGFB or $TGF\beta$) signaling, and PI3K/Akt/mTOR signaling\cite{Ren2020}.
% 13 ARGs (GABARAPL1, ITGA3, USP10, ST13, MAPK9, PRKN, FADD, IKBKB, ITPR1, TP73, MAP2K7, CDKN2A, and EEF2K) with prognostic value were identified in HNSCC patients
%found at Human Protein Atlas dataset (HPA), TCGA, and (GSE6631): A total of 44 patients (22 HNSCC samples and 22 normal samples) were obtained from GSE6631. https://www.proteinatlas.org/ENSG00000103194-USP10/pathology/head+and+neck+cancer

%FOXA2
Prognostic signature integrated of DNA methylation, gene expression, and clinical information could provide a prognostic prediction value for HNSCC patients. Forkhead Box A2 (FOXA2) is significantly associated with the poor prognosis of HNSCC through the studies of CpG-based DNA methylation and RNA expression approach\cite{Shen2017a}.
%32 HNSCC patients had both tumor and adjacent non-tumor tissue samples in TCGA cohort, which was used as the discovery set to identify differential methylation CpG sites.
%  + 0.0256 × cg03774514 FOXA2 => "+" coefficients; higher prognostic score was significantly associated with shorter survival in the training set; 
% https://www.ncbi.nlm.nih.gov/research/pubtator/?view=docsum&query=28852427
%A multi-stage screening strategy => univariate Cox regression was used to evaluate their association with overall survival in the training set, which identified 15 CpG sites with P < 0.05. Further, SIS analysis was performed to further screen out a stable probe combination. Seven of the 15 candidate CpGs were identified, including cg13495205, cg07110405, cg03774514, cg09137696, cg19655456, cg03146625, and cg21546671 (Fig. 2c, Additional file 1: Table S1), mapped to AJAP1, SHANK2, FOXA2, MT1A, ZNF570, HOXC4, and HOXB4, respectively. Using coefficients generated from Cox model, we calculated a prognostic score for each patient based on individualized values of the seven genes (Fig. 2d): prognostic score, methylation = 0.0054 × cg13495205 AJAP1  + 0.0318 × cg07110405 SHANK2 + 0.0256 × cg03774514 FOXA2  + 0.0063 × cg09137696 MT1A  + 0.0013 × cg19655456 ZNF570 − 0.0297 × cg03146625 HOXC4  − 0.0157 × cg21546671 HOXB4 .
%Association between gene expression and methylation. Left panels show correlation of a AJAP1, b SHANK2, c FOXA2, d MT1A, e ZNF570, f HOXC4, and g HOXB4 expression (X-axis) with methylation (Y-axis). Right panels show Kaplan-Meier survival plots of gene expression from the TCGA cohort.

%IL-19 specifically activated an intracellular signal and induced proliferation of the cells, which indicated that IL-19 may act in an autocrine manner in oral cancer.
%IL19 in esophageal squamous cell carcinoma\cite{Hsing2013}, the effects of IL-19 on the esophageal SCC in vivo and in vitro. CE81T cells
%IL-19 may be involved in the pathogenesis of systemic inflammatory diseases. IL-19 is also involved in various inflammatory diseases such as psoriasis, asthma, and rheumatoid arthritis.
% the Chi-Mei Medical Center Institutional Review Board (IRB9705-003). n=60

%%%%%% good guy
%ZNF557 is a tumor suppressor in HPV-positive HNSCC???
%Oncogenic human herpesviruses (e.g., EBV and KSHV) 
%Oncogenic human viruses are silenced through the activities of two members of the Kruppel-associated box (KRAB) domain-zinc finger protein (ZFP) (KRAB-ZFP) epigenetic silencing family, revealing a novel STAT3-KRAB-ZFP axis of virus latency.
%Epstein-Barr virus (EBV) and Kaposi’s sarcoma-associated herpesvirus (KSHV) are silenced by SZF1 and ZNF557, two members of the KRAB-ZFP repressor family\cite{Li2018c}

%calcium/calmodulin-dependent protein kinase II inhibitor (CAMK2N1) is the endogenous inhibitor of CAMK2, which is activated by an increased intracellular Ca2+ concentration.SYT12 plays a critical role in oral cancer and may be a novel therapeutic target. (PMID31598163)
%Overall survival times (as determined by Kaplan-Meier analysis) were markedly longer in patients that had elevated CAMK2N1 expression compared with patients with negative or low CAMK2N1 expression. The miR-182-5p plays a vital role in controlling OSCC cell apoptosis and proliferation by regulating CAMK2N1 expression, which was found to be reduced in OSCC. Down-regulation of miR-182-5p expression attenuated tumor growth, cell survival, and proliferation in vivo through the regulation of the AKT/ERK1/2/NF-κB signaling pathways.
%CAMK2N1 operated as a tumor suppressor gene in patients with HNSCC.
%: 26 results of cancer: ; 1 HNSCC cell line (CAL-27) all said "good guy"
%\cite{Li2018b}
%EMT genes (TCF3, CAMK2N1, EGFR, and IGFBP4) 
%x c) CCLE_RNAseq_rsem_genes_tpm_20180929.txt.gz

\subsection{Limitations of the Work}

Although we combine the power of genome-wide scanning and an optimal cutoff finder for survival analysis, the study has some limitations.
% first limitation:
We are aware of the importance of direct assessment of protein products comprising the basic functional units in cancer cells' biological processes. The announcement of the Cancer Proteome Atlas (\acrshort{tcpa}, http://tcpaportal.org) excites the cancer research community\cite{Li2013c}\cite{Li2017b}. By the utility of the \acrfull{rppa} or \acrfull{rpma}, a microarray kind of "Western blots" in the \acrshort{tcpa} could help to test our hypotheses from \acrshort{rnaseq} study.
However, in the \acrshort{tcpa} database (v3.0\cite{Chen2019}), there are only 237 antibodies available, not covering our candidates so far. %  USP-10 in HPA
%> semi_join(pvalueTex_bad, HPA_bad, by=c("V1"="Gene"))
%validation at protein level:
%At the same time, we found our candidates (DKK1, CAMK2N1, STC2, PGK1, SURF4, NDFIP1, STIP1, DKC1) are also on the list of unfavorable prognostic genes for HNSCC from \acrfull{hpa} (available at https://www.proteinatlas.org/humanproteomepathology/head+and+neck+cancer, accessed 28 September 2020). The ZNF557, ZNF266, and FCGBP are on the list of favorable prognostic genes as well.
% > semi_join(pvalueTex_good, HPA_good, by=c("V1"="Gene"))
%There are 628 genes that consensus of pvalue005KM (6429 genes) and HPA (bad+good in HNSCC)
%%%%%%%%%% modified discussion
%\begin{MyColorPar}{red}

@@

% 其實文章真正的目標是  bioinformatics workflow，而不是 discovered biomarkers itself (see TCGA marker paper\cite{Lawrence2015a})
Moreover, this bioinformatics paper provides targets and supports the community's rationale for looking into these HNSCC candidates by in vitro and in vivo validation. %Collectively, these findings provide new insights into HNSCC and suggest that shared and unique alterations might be leveraged to accelerate progress in prevention and therapy across tumour types\cite{Lawrence2015a}.
%We open invitation for research community by potential support their rationale for looking into these HNSCC candidates.
%邀請科學家們一塊兒來繼續研究下去找出幫助癌症患者的方法
%the rest of the analysis seems fine, but 
%文章的初衷、跨領域：如同deep leaning Keras API，推廣給研究者一個有效的方法 
We aim to promote a reproducible bioinformatics\cite{Preeyanon2014}\cite{Kulkarni2018} workflow allowing successful repetition and extension of analyses based on the TCGA or other in-house HNSCC datasets. % to find more useful biomarkers. % or even other cancer types
% to allow successful repetition and extension of analyses based on original data
A good research reproducibility practice is necessary to allow the reuse of code and results for new projects.
%previously developed methodology to be effectively applied on new data, or to allow 
%reuse of code and results for new projects. %In other words, good habits of reproducibility
It may turn out to be a time-saver in the longer run.
When multiple scientists can reproduce a result, it will also validate our initial results and readiness to progress to the next research phase. 
%why reproducibility?
%Reproducible Bioinformatics Project provides a general schema and an infrastructure to distribute robust and reproducible workflows. Thus, it guarantees to final users the ability to repeat any analysis independently using UNIX-like architecture.
%Utility of this model would facilitate development of more individualized therapy for HNSCC patients and improve prognosis.
Once our laboratory or the community confirms those candidates as the targets, the compound screening\cite{Yang2004}\cite{Hsu2011}\cite{Pathak2021} could facilitate more individualized therapy for HNSCC patients.
%and improve prognosis.
% 1. Pathak, N. et al. Uncovering flexible active site conformations of SARS-COV-2 3Cl proteases through protease pharmacophore clusters and covid-19 drug repurposing. ACS Nano 15, 857–872 (2021). 陽明交大

@@

%Second, GDC data transfer tool (available at https://gdc.cancer.gov/access-data/gdc-data-transfer-tool) allows users to access the open or controlled data. %the GDC data model, data formats
% RESTful API
\end{MyColorPar}

GDC \acrfull{api} is designed under the \acrfull{rest} architecture and provides accessibility to external users for programmatic access to the same functionality found through GDC Portals. Those functions include searching, viewing, submitting, and downloading subsets of data files, metadata, and annotations based on specific parameters. Moreover, if restricted data is requested, the user must provide a token along with the API call. This token can be downloaded directly from the GDC Portals. (paragraph is provided here courtesy of the National Cancer Institute)
% Articles from Cureus are provided here courtesy of Cureus Inc.
Broad GDAC Firebrowse \acrshort{rest}ful \acrshort{api} could be accessed using an R package, FirebrowseR (available at https://github.com/mariodeng/FirebrowseR)\cite{Deng2017}.

>>>>

%%% ***
In summary, identifying predictive biomarkers for selecting standard-of-care or advanced systemic therapy\cite{Cristina2019} in \acrshort{hnscc} is crucial.
Our approach describes a workflow implemented in the R script, which runs on the Rstudio server.
Its function includes data retrieving and pre-processing, feature selection, sliding-window cutoff mining, Kaplan-Meier survival analysis, Cox proportional hazard modeling, and biomarker discovery.
The independent HNSCC datasets were used to validate this strategy.
The workflow, shown in Figure \ref{fig:figure1}, scanned 20,500 human protein-coding genes of the \acrshort{tcga} \acrshort{hnscc} cohort to yield a model with biomarker estimate.

@@

In summary, those three biomarker candidates , clinical T stage, and surgical margin are independent prognosis factors in \acrshort{hnscc}.
We also found those candidates have proper effect size---the Cox's HR either $> 1.8$ or $< 0.6$.
Thus, the prognosis model with coefficients is established from \acrshort{tcga} and GSE65858 \acrshort{hnscc} cohorts. 
%The important input $X_1...X_n$ should be patients' features: age, specific ten gene expressions, clinical T stage, and surgical margin.
%\\[0.5cm]
% total 6+4 =10
% After validation with an independent \acrshort{hnscc} cohort (GSE2837), the result showed six overexpressed genes (symbol as CAMK2N1, PGK1, SURF4, USP10, NDFIP1, FOXA2) are significantly associated with a poor prognosis of overall survival. 
%Furthermore, the four overexpressed genes (symbol as IL19, FCGBP, IQCN - former symbol as KIAA1683, and NPB) are correlated with better survival.

@@

%%% Statistical significance (\textit{P} value) is affected by both sample size, error, effect size (substantive significance)\cite{Sullivan2012}\cite{Thiese2016}, and cutoff.
\subsection{The Purpose of a Sliding-window Cutoff Selection}

% risk of *** Answer 2-3 併入此
%Second, t
Trying to find an optimal cutpoint of that \acrshort{rna} expression data to maximize candidate mining coverage, this strategy could identify more but sometimes weak "biomarkers".
%We have checked this issue in detail (in the Discussions section).
%Thus, the false positives and adjustment of multiple comparisons should be corrected by \acrfull{fdr}.
Thus, we should try our best to handle the effect size from Cox's modeling.
And validation of those candidates is required by using other independent dataset.
% in larger patient populations.% in larger patient
%warrant.

Statistical significance (\textit{P} value) is affected by both sample size, error, effect size (substantive significance)\cite{Sullivan2012}\cite{Thiese2016}, and cutoff. 
The effect size is the magnitude of the difference (e.g., \acrlong{hr}) among comparing groups.
%A \textit{P} value is a tool to help interpret findings from biomedical research.
%In reporting and interpreting studies, both 
%the substantive significance (effect size) and statistical significance (P value) are essential results to be reported.
%Statistical significance depends upon both sample size and effect size. 
The effect size is independent of the sample size\cite{Sullivan2012}.

%When we increase the sample size, the standard error decrease
%, or increase the difference between the sample statistic and hypothesized parameter, the p value decreases, thus making it more likely that we reject the null hypothesis.
In a study with large sample size, the difference could be noticed easily (i.e., \textit{P} value $< 0.05$) due to a decreased standard error\cite{Sullivan2012}.
However, the small effect size (non-zero) is often meaningless or substantive insignificance.
%With a sufficiently large sample, a statistical test will always reveal a significant difference, unless the effect size is exactly zero.
%, that is, when the effect size is exactly zero; 
%yet very small differences, even if significant, are often meaningless. 
%Thus, reporting only the significant P value for an analysis is not adequate for readers to fully understand the results.
%If the magnitude of effect is small and clinically unimportant, the p-value can be "significant" if the sample size is large.
Conversely, the effect size can be large but fail to get statistical significance if the sample size is small. 
%In this scenario, the \textit{P} value tells the effect does not exist.
The following error could also impact the \textit{P} value:
\begin{outline}
\1 A random error, defined as the variability in data, is not considered a bias but rather occurs randomly across the entire study population and can distort the measurement process (e.g., RNA-Seq experiments).
The larger sample size could reduce the random error. 
\1 A systematic error is a bias, which includes selection bias, information bias, and confounding.
It could deleteriously impact the statistical significance.
The larger sample size could not affect the systematic error.
\end{outline}
% effect size is large (whatever is true or artificial), the p value is small
While statistical significance can inform the researcher whether an effect exists, the \textit{P} value will not directly tell the effect size.
Thus, if there is no error in two study groups, and the sample size is the same (not small), the group, which has a larger effect size, will has a small \textit{P} value\cite{Thiese2016}.
%The effect size is the main finding of a quantitative study.
If a skewed cutoff that split, for example, 425 versus 75 between the two groups, it will also get statistical significance by increasing effect size artificially.
% https://stats.stackexchange.com/questions/162359/how-can-you-have-p-0-0001-with-a-sample-size-of-89-and-a-control-group-of-52
%As you can see, even with only 12 events (fustat = 1) that split 10 vs 2 between the two dummy groups you get a log-rank p-value < 0.0001.
%effect size, sample size, cutoff 影響 p value!

@@

We explain aforementioned situation by an example.
%***NDFIP1  % 給他們一個機會
Reviewing the special case of genes, such as NDFIP1, DKC1, PNMA5, and NPB,
we noticed that NDFIP1, with a \textit{P} value of 0.05 at 50\% quantile (median) cutoff, could achieve a \textit{P} value of \num{2.62e-6} at 70\% quantile cutoff.
NDFIP1 has the FDR-adjusted \textit{P} value as \num[round-precision=3, round-mode=figures,
scientific-notation=true]{0.0001074233} (please see supplementary figure S1, a "S"- or "W"-shaped acute-bending on the \textit{P} value plot). % it is \ref{fig:pvaluePlot_NDFIP1
%Moreover, the gene NDFIP1, one of our 20 preliminary candidates, has a \textit{P} value (around $0.05$) at 50\% quantile cutoff, achieving a \textit{P} value of \num{2.62e-6} at 70\% quantile cutoff (please see Figure \ref{fig:pvaluePlot_NDFIP1}, and also in supplementary figure S1).
%After the FDR correction, NDFIP1 still has \textit{P} value of \num[round-precision=3, round-mode=figures, scientific-notation=true]{0.0001074233}.
% the effect size
However, it was excluded as a candidate by small effect size (HR = 1.33 in GSE65858 and HR = 1.42 in GSE2837) less than 1.8.
%%%
\end{MyColorPar} % red

% unclean data: moved from introduction to discussion 
\subsection{Technical Consideration}
\begin{MyColorPar}{red} %%% ***
There are two essential points of biomarker discovery from survival analysis of the TCGA HNSCC dataset.

% dissection
%ZSWIM2\_archive.Rda % 其實最困難的在於 unknown error (ZSWIM2) 
Second, we analyzed the error log during the cutoff finding and Cox modeling, 
the result shows that program running could be halted by several technical problems.
These include:
\begin{outline}

\subsection{Validation by Web-based Tools}

%\subsubsection{HPA} %still TCGA's RNA-Seq
The Human Protein Atlas project (HPA) has proteomics analysis based on 26,941 antibodies targeting 17,165 unique proteins. The HPA's Pathology Atlas analyzes each protein in patients, using \acrshort{ihc} analysis based on tissue microarrays (TMAs) adopted from TCGA. Kaplan-Meier survival analyses are based on RNA-Seq expression levels of human genes in HNSCC tissue and the clinical outcome.
All transcriptomic data were retrieved from the TCGA, and all proteomics were generated in-house using the same antibodies.
The genes (e.g. CAMK2N1, and NDFIP1) are also on the list of unfavorable prognostic genes for HNSCC from \acrshort{hpa} (available at https://www.proteinatlas.org/humanproteome/pathology/head+and+neck+cancer, Version: 20.0 updated: 2020-11-19). 
FCGBP is on the list of favorable prognostic genes as well.
% DKK1, CAMK2N1, STC2, PGK1, SURF4, USP10, NDFIP1, STIP1, and DKC1
% ZNF557, ZNF266, and FCGBP
%Moreover, overexpression of ubiquitin-specific peptidase 10 (USP10), one of the autophagy-related genes, has a poor prognostic impact on HNSCC, which has been proved at mRNA and protein level (HPA database, available at https://www.proteinatlas.org/ENSG00000103194-USP10/pathology/head+and+neck+cancer)\cite{Ren2020}.
% 13 ARGs (GABARAPL1, ITGA3, USP10, ST13, MAPK9, PRKN, FADD, IKBKB, ITPR1, TP73, MAP2K7, CDKN2A, and EEF2K) with prognostic value were identified in HNSCC patients
%found at Human Protein Atlas dataset 
%(Please see Supplementary Figure S3)%\ref{fig:fig_HPA_USP10})
% moved to Supplementary

@@

\subsubsection{Proteomics Validation}

Although we combine the power of genome-wide scanning and an optimal cutoff finder for survival analysis, the study has some limitations.
% first limitation:
We are aware of the importance of direct assessment of protein products comprising the basic functional units in cancer cells' biological processes.
% TCPA, TCIA (immune genes), TCIA (whole-slide image)
The announcement of the Cancer Proteome Atlas (\acrshort{tcpa}, http://tcpaportal.org) excites the cancer research community\cite{Li2013c}\cite{Li2017b}. By the utility of the \acrfull{rppa} or \acrfull{rpma}, a microarray kind of "Western blots" in the \acrshort{tcpa} could help to test our hypotheses from \acrshort{rnaseq} study.
However, in the \acrshort{tcpa} database (v3.0\cite{Chen2019}), there are only 237 antibodies available, not covering our candidates so far. %  USP-10 in HPA
%> semi_join(pvalueTex_bad, HPA_bad, by=c("V1"="Gene"))
%validation at protein level:
%At the same time, we found our candidates (DKK1, CAMK2N1, STC2, PGK1, SURF4, NDFIP1, STIP1, DKC1) are also on the list of unfavorable prognostic genes for HNSCC from \acrfull{hpa} (available at https://www.proteinatlas.org/humanproteomepathology/head+and+neck+cancer, accessed 28 September 2020). The ZNF557, ZNF266, and FCGBP are on the list of favorable prognostic genes as well.
% > semi_join(pvalueTex_good, HPA_good, by=c("V1"="Gene"))
%There are 628 genes that consensus of pvalue005KM (6429 genes) and HPA (bad+good in HNSCC)
%%%%%%%%%% modified discussion
%\begin{MyColorPar}{red}

@@

% 其實文章真正的目標是  bioinformatics workflow，而不是 discovered biomarkers itself (see TCGA marker paper\cite{Lawrence2015a})
Moreover, this bioinformatics paper provides targets and supports the community's rationale for looking into these HNSCC candidates by in vitro and in vivo validation. %Collectively, these findings provide new insights into HNSCC and suggest that shared and unique alterations might be leveraged to accelerate progress in prevention and therapy across tumour types\cite{Lawrence2015a}.
%We open invitation for research community by potential support their rationale for looking into these HNSCC candidates.
%邀請科學家們一塊兒來繼續研究下去找出幫助癌症患者的方法
%the rest of the analysis seems fine, but 
%文章的初衷、跨領域：如同deep leaning Keras API，推廣給研究者一個有效的方法 
We aim to promote a reproducible bioinformatics\cite{Preeyanon2014}\cite{Kulkarni2018} method allowing successful repetition and extension of analyses based on the TCGA or other in-house HNSCC datasets. % to find more useful biomarkers. % or even other cancer types
% to allow successful repetition and extension of analyses based on original data
A good research reproducibility practice is necessary to allow the reuse of code and results for new projects.
%previously developed methodology to be effectively applied on new data, or to allow 
%reuse of code and results for new projects. %In other words, good habits of reproducibility
It may turn out to be a time-saver in the longer run.
When multiple scientists can reproduce a result, it will also validate our initial results and readiness to progress to the next research phase. 
%why reproducibility?
%Reproducible Bioinformatics Project provides a general schema and an infrastructure to distribute robust and reproducible workflows. Thus, it guarantees to final users the ability to repeat any analysis independently using UNIX-like architecture.
%Utility of this model would facilitate development of more individualized therapy for HNSCC patients and improve prognosis.
Once our laboratory or the community confirms those candidates as the targets, the compound screening\cite{Yang2004}\cite{Hsu2011}\cite{Pathak2021} could facilitate more personalized therapy for HNSCC patients.
%and improve prognosis.
% 1. Pathak, N. et al. Uncovering flexible active site conformations of SARS-COV-2 3Cl proteases through protease pharmacophore clusters and covid-19 drug repurposing. ACS Nano 15, 857–872 (2021). 陽明交大

@@

%Second, GDC data transfer tool (available at https://gdc.cancer.gov/access-data/gdc-data-transfer-tool) allows users to access the open or controlled data. %the GDC data model, data formats
% RESTful API
%\end{MyColorPar}

GDC's \acrfull{api} is designed under the \acrfull{rest} architecture and provides accessibility to external users for programmatic access to the same functionality found through GDC Portals. Those functions include searching, viewing, submitting, and downloading subsets of data files, metadata, and annotations based on specific parameters. Moreover, if restricted data is requested, the user must provide a token along with the API call. This token can be downloaded directly from the GDC Portals. (paragraph is provided here courtesy of the National Cancer Institute)
% Articles from Cureus are provided here courtesy of Cureus Inc.
Broad GDAC Firebrowse \acrshort{rest}ful \acrshort{api} could be accessed using an R package, FirebrowseR (available at https://github.com/mariodeng/FirebrowseR)\cite{Deng2017}.
